### 2021-08-23 Frontiers/Frontiers.Rmd
as

---
title: "Palmer amaranth (*Amaranthus palmeri*) adaptation to US Midwest agroecosystems"
runtitle: "Palmer amaranth adaptation"
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Maxwel C Oliveira
    affiliation: '1'
    etal: Oliveira # First author's last name. 
  - name: Amit J Jhala
    affiliation: '2'
  - name: Mark Bernards
    affiliation: '3'
  - name: Chris Proctor
    affiliation: '2'
  - name: Strahinja Stepanovic
    affiliation: '2'
  - name: Rodrigo Werle
    affiliation: '1'
    email: rwerle@uwisc.edu # Indicates corresponding Author
    institution: Department of Agronomy, University of Wisconsin-Madison, United States
    street: 1575 Linden Dr
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    zip: 53705
    country: United States
affiliation:
  - id: '1'
    department: Department of Agronomy
    institution: University of Wisconsin-Madison
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    country: United States
  - id: '2' 
    department: Department of Agronomy and Horticulture
    institution: University of Nebraska-Lincoln
    city: Lincoln
    state: Nebraska # only USA, Australia, Canada
    country: United States
  - id: '3' 
    department: Department of Agronomy
    institution: Western Illinois University
    city: Macomb
    state: Illnois # only USA, Australia, Canada
    country: United States
output: rticles::frontiers_article
csl: frontiers.csl
bibliography: adaptation.bib
---

```{r include=FALSE}
library(kableExtra)
library(tidyverse)
```

\begin{abstract}

Abstract length and content varies depending on article type. Refer to 
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
and length according to article type.

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Evolution Flowering Management Pigweed Weed}

\end{abstract}

# Introduction {-}
Palmer amaranth (*Amaranthus palmeri* S. Watson) is currently ranked as one of the most economically detrimental weed species to cropping systems in the United States [@vanwychen2020]. Unmanaged Palmer amaranth plants compete for water, light, and nutrients, which can drastically impact crop yields [@berger2015]. For example, Palmer amaranth has been documented to reduce up to 91%, 68%, and 54% corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively. Moreover, Palmer amaranth has shown a remarkable capacity to evolve resistance to herbicides. To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017]. Thus, Palmer amaranth poses an economical and environmental risk to sustainable agriculture.

Palmer amaranth is a fast growing summer annual forb indigenous to the Sonoran Desert [@sauer1957]. The species would eventually emerge as a threat to US agriculture in the 1990s. Palmer amaranth weediness is likely a result of human-assisted selection in combination with plant biology. Farm mechanization, adoption of conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013]. On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992]. With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018]. Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017]. The dispersal capacity of Palmer amaranth makes the species one of the most successful cases of weed adaptation to current cropping systems.

Light and temperature are the main environment requirements for Palmer amaranth successful grow and development [@jha2010]. Palmer amaranth is reported with an extended germination period [@ward2013]. Germination of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 C day and night temperatures [@guo2003]. In addition, Palmer amaranth establishment is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010]. Water has not shown to limit Palmer amaranth fitness. Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant-1 [@chahal2018]. Seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021]. Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019]. Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under selection pressure. Palmer amaranth invasion/adaptation into new habitats is likely to increase with the global temperature warming. Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018]. Temperature is a key factor limiting Palmer amaranth expansion to cooler geographies [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018; @briscoerunquist2019].

Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021]. In the United States, Palmer amaranth is well established in the Cotton Belt [@garetson2019; @bagavathiannan2016] in the southern United States but its range is expanding northward. For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a]. There are some reported cases of Palmer amaranth in Michigan [@kohrt2017] and Connecticut [@aulakh2021]. Also, it is estimated that Palmer amaranth can cause high damage to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn make most of US Midwest agronomic hectares. In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021]. Invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021]. Palmer amaranth infestations have not been detected in Canada; however, Palmer amaranth seeds were detected in sweet potato slips in the country [@page2021]. Palmer amaranth is still not as well adapted and established to Northern as it is in the Southern North America. Therefore, its range of expansion into new habitats is increasing. It seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future. Strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed. Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management. Palmer amaranth management should be built on minimizing the species ability to adapt, grow and develop into agroecossystems.

In the southeastern US, early growing Palmer amaranth is well known to have a higher impact on cotton yields compared to late established plants [@macrae2013]. In the northern states, Palmer amaranth impact on the agroecosystem is recent. Studies investigating Palmer amaranth in those locations are limited due to the plant classification as noxious weed species in some northern states [@yu2021]. Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across the northern United States is concerning and warrants investigations on species morphology in such environments. Understanding Palmer amaranth biology and growing strategies under different agroecosystems can enhance our knowledge on species adaptation and management practices. It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and developing resilient and sustainable farming systems [@maclaren2020]. Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.

# Results {-}

<!-- Comments should be HTML-style comments -->

## Plant material and growing conditions {-}

A Palmer amaranth accession (Per1) from Perkins County, Nebraska. with no reported herbicide resistance according to @oliveira2021a was selected for this study. Three weeks prior to the onset of the field experiments, seeds were planted in plastic trays containing potting-mix. Emerged seedlings (1 cm) were transplanted into 200 cm<sup>-3</sup> plastic pots (a plant pot<sup>-1</sup>). Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at the University of Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois University; and kept outdoors in the Perkins extension office at Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.

## Field study {-}

The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N, 'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N, 'W)).

Fields were conventionally tilled prior to crop planting. Corn hybrid and soybean varieties were planted in 76-cm row spacing (Table 1). Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)

```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/weather/Figure 1.png")
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
options(knitr.kable.NA = '')
read_csv("../data/table.csv") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  kable(booktabs = TRUE, caption = "Field study attributes ",
        col.names = c("Attributes",
                           "",
                           "Arlington, WI",
                           "Clay Center, NE",
                           "Grant, NE",
                      "Lincoln, NE", "Macomb, IL")) %>% 
#  pack_rows(
 #   index = c("Bareground" = 2, "Corn" = 4, "Soybean" = 8)) %>% 
  kable_styling(font_size = 10, 
                full_width = FALSE, 
                position = "center",
                latex_options = c("scale_down", "hold_position")) %>% 
  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") 
```

The field experimental units were three adjacent 9.1 m wide (12 rows at 76.2 cm row spacing) by 10.7 m long. The experimental design were arranged in factorial design with three crops, two transplanting times simulating two cohorts, repeated across five locations. Each experimental unit was planted with corn, soybean, or kept under bareground. The two transplanting timings were June 1 (first cohort) and July 1 (second cohort). Palmer amaranth seedlings (potting mix + two seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight plants were equidistantly placed (0.76 m apart) between rows within each crop (Figure 2). After a week, one plant was eliminated and one was kept, resulting in 24 plants per experimental unit and transplanting time (Figure 2). When needed, Palmer amaranth plants were supplied with water during the first week after transplanting.

```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 2.png")
```

After transplanting, Palmer amaranth flowering was monitored until the end of the study. When a plant flowered, the day was recorded, plant sex was identified (male or female), plant height was measured from soil surface to the top of plant. Also, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before weighing (g plant-1 was recorded).

Plants had to be harvested at flowering because Palmer amaranth is neither endemic in Wisconsin nor in Illinois. In our study, all locations follow the methodology of plant harvest at flowering initiation, except in Grant, NE. In this location, all Palmer amaranth plants were harvest at once on July 06, 2018 and 2019 (first cohort), and on August 17, 2018 and on July 31, 2019 (second cohort).

## Statistical analyses {-}

The statistical analyses were performed using R statistical software version 4.0.1.

Analyses of Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015]. Plant height and biomass were log transformed to meet model assumption of normality. In the model, crop (bareground, corn, soybean) was the fixed effect and year nested with location the random effects. Analysis of variance was performed with *anova* function from "car" package [@fox2018]. Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].

The Palmer amaranth flowering timing was estimated as cumulative flowering across all location. Dataset was nested but id did not include Grant, NE. Palmer amaranth cumulative flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].

$$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$$

In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the XXX, and *x* day of year (doy).

The doy for 10, 50, and 90% Palmer amaranth cumulative flowering were determined using the *ED* function of drc package. Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among crops and cohorts using the *EDcomp* function of drc package. The EDcomp function compares the ratio of cumulative flowering using t-statistics, where P-value < 0.05 indicates that we fail to reject the null hypothesis.

Palmer amaranth sex was fitted to a binary logistic regression [@bangdiwala2018]. Binary logistic regression is used for predicting binary classes, such as the probability of a plant being female in a dioecious species. Prior to the analysis, all missing values were removed from the dataset. Also, data from Grant was not used in this analysis due to the uniform plant harvesting at that location. The complete dataset was splitted into 80% train and 20% test data. The 80% train is used for the model training and the 20% test is used for checking how the model generalized on unseen dataset. With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables (without interaction). The model family was binomial with a logit function. The model fit was assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package). The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package. The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function. The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1-score (*f_meas* function) using the "yardstick" package. The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1-score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0. F1-score conveys the balance between the precision and the recall [@yacouby2020].

# Results {-}

<!-- Comments should be HTML-style comments -->

## Palmer amaranth height and biomass {-}

Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (figure 3A). Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and 16.3 g plant-1 in bareground, soybean and corn, respectively. Plants from the second cohort produced 62.6 g plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.

```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 3.png")
```

Palmer amaranth height was more uniform across cohort timings, except when growing in corn (figure 3B). Palmer amaranth plants from the first cohort were on average 69.2 cm tall in bareground, which was not different from the 70.7 cm tall plants from the second cohort timing (P = 0.74). In addition, no difference in Palmer amaranth height (69.3 cm) was detected from first cohort plants in soybean to first and second cohort plants in bareground (P > 0.75). Palmer amaranth plants from the second cohort were nearly 10 cm lower compared to the first cohort in soybeans (P = 0.04). The tallest (first cohort) and smallest (second cohort) Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.

## Palmer amaranth cumulative flowering {-}

Palmer amaranth plants from the first cohort growing in corn resulted in a longer flowering window compared to plants growing in bareground and soybean (Figure 4A). The 10% cumulative Palmer amaranth flowering in soybean, bareground and corn occurred at the end of June. Palmer amaranth reached 10% flowering in soybean, bareground and corn at doy 180, 180.9 and 181.7, respectively. The 50% Palmer amaranth cumulative flowering occurred in July. Palmer amaranth reached 50% flowering in bareground, soybean and corn at doy 193.4, 194.8, and 206.6, respectively. Similar trend was observed at 90% Palmer amaranth cumulative flowering. Palmer amaranth reached 90% flowering at doy 252.6 in corn (early September), which was 38 and 32 days after reaching 90% flowering in bareground and soybean, respectively.

```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 4.png")
```

Palmer amaranth cumulative flowering at the second cohort ranged from mid July to mid September (Figure 4B). Palmer amaranth growing in the bareground resulted in earlier flowering time compared to soybean and corn. Palmer amaranth growing in bareground reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively. Palmer amaranth growing in soybean reached 10% flowering at doy 210.9, which was 6 days prior to corn (*P*-value = 0.00). Similar trend was observed at 50% flowering, whereas Palmer amaranth reached 50% flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; *P* = 0.00). The 90% Palmer amaranth cumulative flowering occurred at same day in corn (260.9) and soybean (260.5; *P* = 0.66).

## Palmer amaranth gender {-}

The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively. The likelihood ratio test showed a p-value of < 0.00. The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.8% less females plants (Table 2). Moreover, increasing a unit doy  increases the probability of having a female plant by 0.4% (Table 2 and Figure4A). Similar trend is observed for weight as well as height, whereas the probability of being female increase by 0.2% (Figure 4B) and 0.1% (Figure 4C) when a unit of weight (g) and height (cm) increases, respectively.

```{r echo = FALSE, warning = FALSE, message=FALSE}
read_csv("../data/table_a.csv") -> table_a

#names(table_a)[2] <- paste0(names(table_a)[2], 
#                                footnote_marker_symbol(1))

table_a %>% 
  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.") %>% 
  kable_styling(font_size = 10, 
                full_width = FALSE, 
                position = "center",
                latex_options = c("hold_position")) %>% 
  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") %>%
  footnote(
           alphabet = c("Average Marginal Effects.", "Standard Error."),
           footnote_as_chunk = TRUE)
```

```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 5.png")
```

The model accuracy evaluation accuracy in the 20% test dataset was 0.62 with a cutoff value for female and male plants of 0.43. The model classification showed a precision of 0.64, recall of 0.66, and a F means score of 0.65. In addition, the area under the curve was 0.64.

# Discussion {-}

Our study showed that Palmer amaranth biomass, height, flowering pattern and gender varied within agroecossystems and cohort timings.  In general, first cohort of Palmer amaranth plants were heavier and taller when compared to the second cohort. At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for the species. High biomass and taller Palmer amaranth plants are likely a weed strategy to compete for light in between crop rows in absence of canopy. In such conditions, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem. This is evident when comparing Palmer amaranth canopy shape, and its extended flowering pattern when growing into corn compared to soybean. The Palmer amaranth competition strategy was to mimic the crop grow and development (Figure 6). These results suggests that Palmer amaranth can fast evolve life-history traits to adapt into cropping systems and cultural practices, which was also showed in a study varying nitrogen fertilization [@bravo2018]. Our results highlight the Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].

```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 6.png")
```

Palmer amaranth grow and development in second cohort was limited due to the crop competitive ability at advanced development stages. Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competitiveness. As a result, Palmer amaranth height and biomass was lower compared to its first cohort. Under crop canopy (e.g., second cohort), Palmer amaranth flowering window was near to similar in corn and soybean. Palmer amaranth growing without crop competition produced the highest amounts of biomass and less extended flowering window. The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height. Nonetheless, Palmer amaranth produced 21% less biomass in second cohort compared to first cohort timing. In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015]. These results suggest that crop competition is not the only factor limiting late Palmer amaranth establishment. The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to day length, light availability and thermal units (e.g, growing degree days). It is hypothesize that reduced day length contributed to smaller plants at second cohort as well as shorter flowering period. A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021]. In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established near to crop planting.

Seed production was not evaluated due to plant harvest at initiation of flowering. Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018]. I our study, Palmer amaranth growing at first cohort accumulated an overall 36% more biomass when compared to second cohort. Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the cropping system. Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015]. Still, seed production at second cohort is likely to replenish the soil seedbank. Seed production and deposition in the seedbank is also a key factor for species perpetuation [@menges1987]. Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013]. Therefore, preventing Palmer amaranth seed production or/and seed migration to its non-native habitat is an essential strategy to minimize weed impact into agroecossystem [@davis2015].

An ecological approach to reduce seed production in Palmer amaranth is understanding its flowering window. Our study suggests that Palmer amaranth flowering was slightly influenced by cropping systems and cohort timings. Palmer amaranth growing in bareground and corn resulted in the overall shortest and longest flowering window, respectively. When growing in soybean, Palmer amaranth flowering window was similar to bareground at first cohort but similar to corn at second cohort timing. Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965]. We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window. Early flower initiation is plant strategy when growing in stress conditions. For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021]. A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, and thus reduce plant seed production and exchange of resistant alleles [@jhala2021]. Sex dimorphism manipulation is considered a potential ecological pest control [@mcfarlane2018; @schliekelman2005].

The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888]. In our study, the gender model performance was decent (AIC 0.64) considering the biology of plant flowering. A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930]. However, a slight deviation from 1:1 sex ratio might occur in some dioecious species. For example, the dioecious *Halophila stipulacea* is a female-biased plant in its native habitat, but the naturalized *H. stipulacea* have a 1:1 ratio [@nguyen2018]. Naturalized of *H. stipulacea* reduced female-male ratio to expand into its non-native habitat [@nguyen2018]. Also, biotic and/or abiotic stress can influence plant sex determination. Palmer amaranth  male-to-female ratio was greater under high plant densities [@korres2017a] and after herbicide application [@rumpa2019]. Our model estimated that late flowering, heavier and taller Palmer amaranth plants deviated from 1:1 ratio in favor to female plants. It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017]. Our model also estimated more female plants in soybean and bareground compared to corn, which might linked to plant competition strategy in each agroecossystem. Our results showed the influence of life-history and ecological traits on sexual dimorphism in Palmer amaranth. Sexual dimorphism is documented in other dioecious species [@barrett2013]. For example, stronger female plant competition and greater male tolerance to herbivory was reported in *Spinacia oleracea* [@perez-llorca2019]. Research on candidate genes for sex determination in *Amaranthus* species are currently underway but it is far to complete [@montgomery2021; @montgomery2019]. Further studies are also needed to understand the ecological basis of Palmer amaranth flowering, including the plant behavior under climate change.

Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems. Is likely that Palmer amaranth range will continue to expanding to new geographies. Therefore, preventive management is a priority to minimizing Palmer amaranth dispersal. Reactive management should focus on early-season management programs, which would have a large negative effect on Palmer amaranth growth and development. Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation [@oliveira2021a], plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g. cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems. The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on economy and sustainability.

# Disclosure/Conflict-of-Interest Statement {-}

<!--  
Frontiers follows the recommendations by the International Committee of Medical
Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
that all financial, commercial or other relationships that might be perceived by
the academic community as representing a potential conflict of interest must be
disclosed. If no such relationship exists, authors will be asked to declare that
the research was conducted in the absence of any commercial or financial
relationships that could be construed as a potential conflict of interest. When
disclosing the potential conflict of interest, the authors need to address the
following points:

 - Did you or your institution at any time receive payment or services from a
   third party for any aspect of the submitted work?
 - Please declare financial relationships with entities that could be perceived
   to influence, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 - Please declare patents and copyrights, whether pending, issued, licensed
   and/or receiving royalties relevant to the work.
 - Please state other relationships or activities that readers could perceive to
   have influenced, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 -->

The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

# Author Contributions {-}

<!--

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research. All authors reviewed the manuscript.

# Acknowledgments

Funding: This work received no specific grant from any funding agency, commercial, or not-for-profit sectors

# Supplemental Data

Supplementary Material should be uploaded separately on submission, if there are
Supplementary Figures, please include the caption in the same file as the
figure. LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

# References

A Frontier article expect the reference list to be included in this section.
To make that happens, the below syntax can be used. This [feature is from Pandoc citeproc](https://pandoc.org/MANUAL.html#placement-of-the-bibliography) which is used with `frontier_article()` to handle the bibliography

::: {#refs}
:::

# Figures

```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "160mm", out.height = "100mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit', out.width = "150mm", out.height = "70mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B). From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```