### 2021-08-23 data analysis/sex_analysis.Rmd

---
title: "sex"
author: "Maxwel Coura Oliveira"
date: "8/11/2021"
output: html_document
---

```{r}
library(emoGG)
library(tidyverse)
library(lubridate)
library(tidymodels)
library(yardstick)
library(broom)       # for making model summary tidy
library(visreg)      # for potting logodds and probability 
library(margins)     # to calculate Average Marginal Effects
library(rcompanion)  # to calculate pseudo R2
library(ROCR) 
#https://towardsdatascience.com/modelling-binary-logistic-regression-using-tidymodels-library-in-r-part-1-c1bdce0ac055
#https://towardsdatascience.com/modelling-binary-logistic-regression-using-r-research-oriented-modelling-and-interpretation-f67b3a954101
```

```{r}
read_csv("../data/data2.csv") %>% 
#  filter(location != "grant") %>% 
  mutate(location = str_replace(location, "havelock", "lincoln")) %>% 
  mutate(harvest = case_when(
    location == "clay center" & doyh == 169 ~ "8/18/2018",
    TRUE ~ harvest
  )) %>% 
  janitor::clean_names() %>% 
  rename(planting = planting_4,
         planting_1 = planting_9) -> data1
```

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting),
         doyh = yday(harvest)) %>% 
  mutate_if(is_character, as_factor) %>% 
  filter(!is.na(gender)) %>% 
#  filter(!is.na(weight)) %>% 
  mutate(fem = as_factor(if_else(gender == "f", 1, 0))) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) %>% 
  filter(location != "grant") %>% 
  dplyr::select(fem, gender, location, year, month, crop, height, weight, doyh) -> data2
  na.omit(data2) -> data2
```

```{r}
data2 %>% 
  distinct(location)
```

```{r}
# Total number of rows in the data frame
n <- nrow(data2)
# Number of rows for the training set (80% of the dataset)
n_train <- round(0.80 * n)
# Create a vector of indices which is an 80% random sample
set.seed(123)
train_indices <- sample(1:n, n_train)
# Subset the data frame to training indices only
train <- data2[train_indices, ]
# Exclude the training indices to create the test set
test <- data2[-train_indices, ]
```

```{r}
paste("train sample size: ", dim(train)[1])
paste("test sample size: ", dim(test)[1])
```

```{r}
library(recipes)
recipe(fem ~ doyh + crop + height + weight, data = train) %>% 
  recipes::step_dummy(all_nominal(), -all_outcomes()) -> recip1

#preprocessing
prepped_rec <- prep(recip1, verbose = TRUE, retain = TRUE )

# new data set with juice
juice <- juice(prepped_rec)

# bake to see the preprocessed training data
preproc_train <- recipes::bake(prepped_rec, new_data = NULL)
glimpse(preproc_train)
# bake to see the preprocessed testing data
preproc_test <- recipes::bake(prepped_rec, new_data = test)
glimpse(preproc_test)
```

```{r}
#Fitting a binary logistic regression
model_logi <- glm(fem ~ ., data = preproc_train, family = "binomial")
#Model summary
summary(model_logi)
```

```{r}
# Pseudo R_squared values and Likelyhood ratio test
rcompanion::nagelkerke(model_logi)
```

```{r}
# Probabilities of female with doy
visreg(model_logi, "doyh", scale="response", rug=2, gg=TRUE,
       xlab="Day of year",
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> doy_f
```

```{r}
doy_f +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(170, 270), breaks = seq(170, 270, 20)) +
  theme_test() +
  geom_emoji(aes(x = 220, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") -> doyf

#ggsave("figures/fig2.png")
```

```{r}
# Probabilities of female with weight
visreg(model_logi, "weight", scale="response", rug=2, gg = TRUE,
       xlab=expression(paste("Weight (g plant "^"-1",")")),
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> weight_f
```
```{r}
weight_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  scale_x_continuous(limits = c(0, 1100), breaks = seq(0, 1100, 150)) +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> weightf

#ggsave("figures/fig1.png")
```

```{r}
# Probabilities of female with weight
visreg(model_logi, "height", scale="response", rug=2, gg = TRUE,
       xlab=expression(paste("Height (cm plant "^"-1",")")),
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> height_f
height_f
```

```{r}
height_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> heightf

#ggsave("figures/fig.png")
```

```{r}
library(patchwork)

doyf + weightf + heightf +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, hjust = 0, vjust = 0))

ggsave("figures/Figure 5.png", width = 9, height = 3)
```