### 2022-02-18 Manuscript/Manuscript.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass[utf8]{frontiersSCNS}

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% BELOW TAKEN FROM rticles plos template
%
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Below is from frontiers
%
\bibliographystyle{frontiersinSCNS}
% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\def\Authors{
  Maxwel C. Oliveira\,\textsuperscript{1},
  Amit J. Jhala\,\textsuperscript{2},
  Mark L. Bernards\,\textsuperscript{3},
  Chris Proctor\,\textsuperscript{2},
  Strahinja Stepanovic\,\textsuperscript{2},
  Rodrigo Werle\,\textsuperscript{1*}}

\def\Address{

  \textsuperscript{1} Department of Agronomy, University of
Wisconsin-Madison,  Madison,  Wisconsin,  United States

  \textsuperscript{2} Department of Agronomy and
Horticulture, University of
Nebraska-Lincoln,  Lincoln,  Nebraska,  United States

  \textsuperscript{3} School of Agriculture, Western Illinois
University,  Macomb,  Illinois,  United States
  }

  \def\firstAuthorLast{Oliveira {et~al.}}

  \def\corrAuthor{Rodrigo Werle}\def\corrAddress{Department of Agronomy,
University of Wisconsin-Madison, United States\\1575 Linden
Dr\\Madison, Wisconsin, 53705 United
States}\def\corrEmail{\href{mailto:rwerle@uwisc.edu}{\nolinkurl{rwerle@uwisc.edu}}}

\begin{document}
\onecolumn
\firstpage{1}

\title[Palmer amaranth adaptation]{Palmer amaranth (\emph{Amaranthus
palmeri}) adaptation to US Midwest agroecosystems}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is one of the most troublesome agronomic weed species in the United States. Palmer amaranth is prevalent in the Southern Great Plains and Southeastern United States, and its range is expanding northward through natural dispersal and human intervention. Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies. A study was conducted to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations in the US Midwest. Results demonstrated that the first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from the second cohort (established in July). The first Palmer amaranth cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively. Palmer amaranth height was most impacted when growing in corn, and averaged 85.2 cm tall in the first cohort, and 38.2 cm tall in the second cohort in corn. Moreover, Palmer amaranth flowering window shifted according to crop and cohort timings. Palmer amaranth growing in intense competition, such as under low light in corn, resulted in the longest flowering window. Palmer amaranth gender was slightly influenced by day of year, weight and height. We documented a high degree of plasticity in Palmer amaranth, which will presumably favor its adaptation and expansion in cropping systems north of its current range. Therefore, preventing Palmer amaranth dispersal into new habitats is the most effective management strategy.

\tiny
 \keyFont{ \section{Keywords:} Evolution, Flowering, Management, Pigweed, Weed}

\end{abstract}

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
ranked as one of the most economically detrimental weed species to
cropping systems in the United States (Van Wychen, 2020). Unmanaged
Palmer amaranth plants compete for water, light, and nutrients, which
can drastically impact crop yields (Berger et al., 2015). For example,
Palmer amaranth has been documented to reduce up to 91\%, 68\%, and 54\%
corn (Massinga et al., 2001), soybean (Klingaman and Oliver, 1994), and
cotton (Morgan et al., 2001) yields, respectively. Moreover, Palmer
amaranth has shown a remarkable capacity to evolve resistance to
herbicides. To date, Palmer amaranth has evolved resistance to eight
herbicide sites of action (Heap, 2021), increasing the weed management
complexity (Lindsay et al., 2017) and posing an economical and
ecological risk to row-crop agriculture.

Palmer amaranth as a problem weed is a function of both inherent
adaptations and selected management practices. Palmer amaranth is a fast
growing summer annual forb indigenous to the Sonoran Desert (Sauer,
1957). It became a serious problem weed in US agriculture in the 1990s
(Ward et al., 2013). Palmer amaranth weediness is likely a result of
human-assisted selection combined with plant biology. Farm
mechanization, adoption of conservation agriculture (e.g., no-till), and
intensive use of herbicides for weed management are the main
human-mediated selections of Palmer amaranth in cropping systems (Ward
et al., 2013). Palmer amaranth is a prolific seed producer with a C4
photosynthetic apparatus (Wang et al., 1992). With dioecious nature,
Palmer amaranth male and female plants are obligate outcrossers,
increasing the chances of exchanging adaptive traits among plants (Jhala
et al., 2021; Oliveira et al., 2018). Also, Palmer amaranth's small
seeds (e.g, 1 mm) tend to thrive in no-tillage systems (Price et al.,
2011), and spread across locations through farm equipment (Sauer, 1972),
seed mixes (Hartzler and Anderson, 2016), wildlife (Farmer et al.,
2017), etc., making it one of the most successful examples of weed
adaptation to current cropping systems.

Palmer amaranth's plasticity allows it to respond successfully to
environmental changes. Palmer amaranth demonstrates a high degree of
plasticity to light, temperature, water availability, and human
management (Jha et al., 2010). Palmer amaranth has an extended
germination period of up to \_\_\_\_\_ (Ward et al., 2013). Germination
of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth
(Keeley et al., 1987), and optimal germination and biomass production
occurred at 35/30 C day and night temperatures (Guo and Al-Khatib,
2003). Palmer amaranth emergence can be influenced by tillage and the
use of preemergence herbicides (Chahal et al., 2021), and may result in
weed germination shifts within a population, as documented in
\emph{Bassia scoparia} (Sbatella and Wilson, 2010). In an experiment
where Palmer amaranth was subjected to continuous water stress, it
survived and produced at least 14000 seeds plant\textsuperscript{-1}
(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
water conditions were heavier, less dormant, and prompt for germination
(Matzrafi et al., 2021). Growing conditions and management practices
also influence Palmer amaranth sex dimorphism and flowering pattern
(Korres et al., 2017; Rumpa et al., 2019).

Palmer amaranth's current global range includes agronomic cropland in
Italy (Milani et al., 2021), soybean producing regions of Brazil and
Argentina (Larran et al., 2017; Küpper et al., 2017), and the Cotton
Belt (Garetson et al., 2019; Bagavathiannan and Norsworthy, 2016) and
southern Great Plains of the United States (Kumar et al. (2020), Crespo
et al. (2016){]}. However, its range appears to be expanding steadily,
and in some cases rapidly, in the U.S. For example, in states with large
Palmer amaranth infestations its range is gradually pushing north as it
displaces other weeds in agronomic row crops (R Werle, personal
observation), aided by wildlife (Farmer et al., 2017) and agricultural
machinery. But there are other examples where it has suddenly become a
problem weed hundreds of miles away from known infestations. New
infestations in Michigan, Indiana and North Dakota appear to have
resulted from Palmer amaranth contaminated livestock feed. New
infestations in Iowa and Minnesota were associated with planting
contaminated pollinator habitat seed (Yu et al., 2021). Palmer amaranth
infestations have not been detected in Canada, but Palmer amaranth seeds
were found in sweet potato slips imported into the country (Page et al.,
2021). Global warming will create new opportunities for Palmer amaranth
invasion. Although agronomic crops in warm environments like Australia
and Sub-Saharan Africa are currently at greatest risk for Palmer
amaranth invasion, warming temperatures will reduce barriers that may
have limited Palmer amaranths spread into cooler climates like Canada
and Northern Europe (Kistner and Hatfield, 2018; Briscoe Runquist et
al., 2019).

There are many areas in the U.S. Corn Belt (North Central States) where
Palmer amaranth is not yet established, and its potential adaptability
is untested. Nonetheless, the rapid expansion of Palmer amaranth across
the Northern U.S. is concerning and warrants investigations on its
adaptability. We do know that Palmer amaranth caused yield loss in
Illinois soybean fields (Davis et al., 2015), but an Iowa study showed
that Palmer amaranth was not as well adapted as waterhemp
(\emph{Amaranthus tuberculatus}) to conditions in that geography (Baker,
2021). We also know that Palmer amaranth plants that establish shortly
after row-crop planting have a much greater impact on crop yield than
plants that emerge after the crop has produced several leaves (MacRae et
al., 2013). Understanding Palmer amaranth morphology and development
under different agroecosystems and across a wide geography can enhance
our knowledge of its adaptability, and may also aid in designing
effective tactics to limit is range expansion and minimize its negative
effects on row crops. The objective of this study was to investigate the
flowering pattern, gender, biomass production, and height of Palmer
amaranth cohorts growing in corn, soybean and bareground environments
across five locations in the North Central United States.

\hypertarget{material-and-methods}{%
\section*{Material and Methods}\label{material-and-methods}}
\addcontentsline{toc}{section}{Material and Methods}

\hypertarget{plant-material-and-growing-conditions}{%
\subsection*{Plant material and growing
conditions}\label{plant-material-and-growing-conditions}}
\addcontentsline{toc}{subsection}{Plant material and growing conditions}

A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no
glyphosate resistance was selected for this study (Oliveira et al.,
2021). Three weeks prior to the establishment of each cohort, seeds were
planted in plastic trays containing potting-mix. Emerged seedlings (1
cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (a
plant pot\textsuperscript{-1}). Palmer amaranth seedlings were supplied
with adequate water and kept under greenhouse conditions at the
University of Wisconsin-Madison, University of Nebraska-Lincoln, and
Western Illinois University; and kept outdoors at the Perkins extension
office in Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when
they were transported to the field.

\hypertarget{field-study}{%
\subsection*{Field study}\label{field-study}}
\addcontentsline{toc}{subsection}{Field study}

The experiment was conducted in 2018 and 2019 under field conditions at
five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
(40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N,
96.42'W), and Macomb, IL (XXX'N, XXX'W).

Fields were conventionally tilled prior to crop planting. Corn and
soybean were planted in 76-cm row spacing (Table 1). Monthly mean air
temperature and total precipitation were obtained using Daymet weather
data from June through September across the five locations in 2018 and
2019 (Correndo et al., 2021) (Figure 1).

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-2}Field study attributes }
\centering
\resizebox{\linewidth}{!}{
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lllllll}
\toprule
Attributes &  & Arlington, WI & Clay Center, NE & Grant, NE & Lincoln, NE & Macomb, IL\\
\midrule
Bareground &  &  &  &  &  & \\

 & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\

Corn &  &  &  &  &  & \\

 & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\

 & Seeding rate & 88956 & 86487 &  &  & \\

 & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\

 & Stage at 1 cohort & V2-3 &  &  &  & \\

 & Stage at 2 cohort & V6-7 &  &  &  & \\

 & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\

\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\

Soybean &  &  &  &  &  & \\

 & Variety & DSR-1950 & AG21X8 &  &  & \\

 & Seeding rate & 296400 & 321237 &  &  & \\

 & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\

 & Stage at 1 cohort & V1-2 &  &  &  & \\

 & Stage at 2 cohort & V5-6 &  &  &  & \\

\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\

Soil &  &  &  &  &  & \\

 & Type & Plano-silt-loam & Crete Silt Loam &  &  & \\

 & Ratio (sand-clay-silt) & 10-64-26 & 17-58-25 &  &  & \\

 & pH & 6.6 & 6.5 &  &  & \\

\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
\end{tabular}}
\end{table}

The field experimental units were three adjacent 9.1 m wide (12 rows at
76.2 cm row spacing) by 10.7 m long. The experimental design were
arranged in factorial design with three crops, two transplanting times
simulating two cohorts, repeated across five locations. Each field
experimental unit was planted with corn, soybean, or maintained as
bareground. The two transplant timings were June 1 (first cohort) and
July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight
plants were equidistantly placed (0.76 m apart) between rows within each
crop. After a week, one plant was eliminated and one was kept, resulting
in 24 plants per experimental unit and transplanting time (Figure 2).
When needed, Palmer amaranth plants were supplied with water during the
first week after transplanting to assure seedling survival.

After transplanting, Palmer amaranth flowering was monitored until the
end of the study. When a plant flowered, the day was recorded, plant
gender was identified (male or female), plant height was measured from
soil surface to the top of plant. Also, aboveground plant, and all
aboveground plant organs were harvested, then oven dried at 65 C until a
constant weight was reached, and dry biomass (g
plant\textsuperscript{-1}) was recorded. Plants were harvested at
flowering because Palmer amaranth infestations do not occur at the
Wisconsin or Illinois research locations. In our study, all locations
followed the methodology of plant harvest at flowering initiation,
except in Grant, NE. At the Grant, NE, location, all plants from the
first cohort were harvested on July 6, 2018 or July 6, 2019, and all
plants from the second cohort were harvested on August 17, 2018 or July
31, 2019, regardless of flowering status.

\hypertarget{statistical-analyses}{%
\subsection*{Statistical analyses}\label{statistical-analyses}}
\addcontentsline{toc}{subsection}{Statistical analyses}

The statistical analyses were performed using R statistical software
version 4.0.1 (Team, 2021).

Analyses of Palmer amaranth height and biomass were performed with a
linear mixed model using \emph{lmer} function from ``lme4'' package
(Bates et al., 2015). Plant height and biomass were log transformed to
meet model assumption of normality. In the model, crop (bareground,
corn, soybean) and cohort time (first and second) were the fixed effects
and year nested with location were the random effects. Analysis of
variance at \(\alpha\) 0.05 was performed with \emph{anova} function
from ``car'' package (Fox and Weisberg, 2018). Marginal means and
compact letter display were estimated with \emph{emmeans} and \emph{cld}
from packages ``emmeans'' (Lenth et al., 2021) and ``multcomp'' (Hothorn
et al., 2008), respectively.

Palmer amaranth cumulative flowering estimation was determined across
all locations, except Grant, NE. Cumulative flowering estimation was
determined using an asymmetrical three parameter log logistic Weibull
model of the drc package (Ritz et al., 2015):

\[Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\]

In this model, \emph{Y} is the cumulative flowering, \emph{d} is the
upper limit (set to 100), and \emph{e} is the inflection point, and
\emph{x} is the day of year (doy).

The doy for 10, 50, and 90\% cumulative flowering were determined using
the \emph{ED} function of drc package. Also, the 10, 50, and 90\% Palmer
amaranth cumulative flowering were compared among crops and cohorts
using the \emph{EDcomp} function of drc package. The EDcomp function
compares the ratio of cumulative flowering using t-statistics, where
P-value \textless{} 0.05 indicates that we fail to reject the null
hypothesis.

A binary logistic regression was fitted to Palmer amaranth gender.
Binary logistic regression is used for predicting binary classes
(Bangdiwala, 2018), such as the probability of a plant being female in a
dioecious species. Prior to the analysis, missing values were removed
from the dataset (including all data from the Grant location). The
resulting dataset was split into 80\% train and 20\% test data. The 80\%
train is used for model training and the 20\% test is used for checking
model performance on unseen dataset. Using the 80\% train data, a
generalized linear model (base R \emph{glm} function) was fitted to
binary response variable, the probability of being female (0 to male and
1 to female). The independent variables were harvest day of year,
height, weight, and crop (without interaction). The model family was
binomial with a logit function. The model fit was assessed through
pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and
likelihood ratio using \emph{nagelkerke} function from ``rcompanion''
package (Mangiafico, 2021). The marginal effects computation was
performed with Average Marginal Effects (AMEs) at every observed value
of x and averaged across the results (Leeper, 2017) using \emph{margins}
function from ``margins'' package (Leeper et al., 2021). The 20\% test
data was predicted using the \emph{predict} function with a cutoff
estimation for male or female using \emph{performance} function from
ROCR package (Sing et al., 2005). The model quality prediction from the
classification algorithm was measured with precision (\emph{precision}
function), recall (\emph{recall} function) and F1-score (\emph{f\_meas}
function) using the ``yardstick'' package (Kuhn et al., 2021). The
precision determines the accuracy of positive predictions (female
plants), recall determines the fraction of positives that were correctly
identified, and F1-score is a weighted harmonic mean of precision and
recall with the best score of 1 and the worst score of 0 (Raoniar,
2021). F1-score conveys the balance between the precision and the recall
(Yacouby and Axman, 2020). The area under the receiver operating curve
(AUC-ROC) was also estimated with performance function using the true
positive and false positive rates. The higher the AUC, better the model
is at distinguishing between female and male Palmer amaranth.

\hypertarget{results}{%
\section*{Results}\label{results}}
\addcontentsline{toc}{section}{Results}

\hypertarget{palmer-amaranth-height-and-biomass}{%
\subsection*{Palmer amaranth height and
biomass}\label{palmer-amaranth-height-and-biomass}}
\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}

Palmer amaranth plants accumulated more biomass when growing in
bareground compared to plants growing in soybean and corn (figure 3A).
Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3
g plant\textsuperscript{-1} in bareground, soybean and corn,
respectively. Plants from the second cohort produced 62.6 g
plant\textsuperscript{-1} in bareground, followed by 6.3 g
plant\textsuperscript{-1} in soybean, and 1.4 g
plant\textsuperscript{-1} in corn.

Palmer amaranth height was less affected by cohort timing than was
weight, with the exception of plants growing in corn (Figure 3B). Plants
from the first cohort were on average 69.2 cm tall in bareground, which
was not different from the 70.7 cm tall plants from the second cohort
timing (P = 0.74). In addition, no difference in Palmer amaranth height
(69.3 cm) was detected from first cohort plants in soybean to first and
second cohort plants in bareground (P \textgreater{} 0.75). Palmer
amaranth plants from the second cohort were nearly 10 cm shorter
compared to the first cohort in soybeans (P = 0.04). The tallest (85.2
cm, first cohort) and shortest (38.2 cm, second cohort) plants were
observed in corn.

\hypertarget{palmer-amaranth-cumulative-flowering}{%
\subsection*{Palmer amaranth cumulative
flowering}\label{palmer-amaranth-cumulative-flowering}}
\addcontentsline{toc}{subsection}{Palmer amaranth cumulative flowering}

The initiation and duration of Palmer amaranth flowering was strongly
influenced by cohort and surrounding vegetation (Figure 4A, 4B). In the
first cohort, floral initiation (10\% flowering) occurred near the end
of June for all three treatments, at doy 180, 180.9, and 181.7 for
soybean, bareground, and corn, respectively. In the second cohort,
floral initiation occurred earlier on plants growing in the bareground
than plants growing in soybean or corn (doy 203.8 versus doy 210.9 or
216.8, respectively). Palmer amaranth growing in the bareground had the
shortest duration of floral initiation (measured as the difference
between 10\% and 90\% cumulative flowering) in both the first (34 days)
and second (28 days) cohorts. Palmer amaranth growing in soybean had a
shorter duration of flowering than corn in the first cohort (40 days vs
71 days), but a longer duration in the second cohort (50 vs 44 days).

Palmer amaranth cumulative flowering in the second cohort ranged from
mid July to mid September (Figure 4B). Palmer amaranth growing in the
bareground resulted in earlier flowering time compared to soybean and
corn. Palmer amaranth growing in bareground reached 10\%, 50\%, and 90\%
flowering time at day 203.8, 214.4, and 232.2, respectively. Palmer
amaranth growing in soybean reached 10\% flowering at doy 210.9, which
was 6 days prior to corn (\emph{P}-value = 0.00). Similar trend was
observed at 50\% flowering, whereas Palmer amaranth reached 50\%
flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; \emph{P}
= 0.00). The 90\% Palmer amaranth cumulative flowering occurred at same
day in corn (260.9) and soybean (260.5; \emph{P} = 0.66).

\hypertarget{palmer-amaranth-gender}{%
\subsection*{Palmer amaranth gender}\label{palmer-amaranth-gender}}
\addcontentsline{toc}{subsection}{Palmer amaranth gender}

The model goodness of fit was 0.23, 0.32, 0.40 using pseudo R-squared
test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
The likelihood ratio test showed a p-value of \textless{} 0.00. The
average marginal effects showed that Palmer amaranth growing in corn
resulted in 14.8\% fewer females plants (Table 2). Moreover, increasing
a cumulative flowering unit doy increases the probability of having a
female plant by 0.4\% (Table 2 and Figure5A). A similar trend was
observed for weight (Figure 5B) and height (Figure 5C), where the
probability of being female increased 0.1\% and 0.2\% for each unit
increase of weight (g) and height (cm), respectively (Table 2).

\textbackslash begin\{table\}{[}!h{]}

\textbackslash caption\{\label{tab:unnamed-chunk-3}Average marginal
effects of Palmer amaranth gender logistic model. Factor pararemter
values (crop\_corn and 'crop\_bareground) are calculated relative to
soybean.\} \centering \fontsize{10}{12}\selectfont

\begin{tabular}[t]{lrrrrrr}
\toprule
Term & AME & SE & Lower & Upper & Z-score & P-value\\
\midrule
crop\_bareground & -0.048 & 0.054 & -0.154 & 0.059 & -0.876 & 0.381\\

crop\_corn & -0.148 & 0.052 & -0.250 & -0.046 & -2.842 & 0.004\\

doyh & 0.004 & 0.001 & 0.003 & 0.006 & 4.959 & 0.000\\

height & 0.002 & 0.001 & 0.001 & 0.003 & 2.953 & 0.003\\

weight & 0.001 & 0.000 & 0.000 & 0.001 & 2.179 & 0.029\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} Average Marginal Effects. \textsuperscript{b} Standard Error.}\\
\end{tabular}

\textbackslash end\{table\}

The model accuracy evaluation accuracy in the 20\% test dataset was 0.62
with a cutoff value for female and male plants of 0.43. The model
classification showed a precision of 0.64, recall of 0.66, and a
F1-score of 0.65. In addition, the AUC was 0.64.

\hypertarget{discussion}{%
\section*{Discussion}\label{discussion}}
\addcontentsline{toc}{section}{Discussion}

Our study confirmed that Palmer amaranth is well adapted to growing
throughout the Midwestern United States, and is simply limited in its
range by seed dispersal (Davis et al.~2015). We also confirmed Palmer
amaranth's extraordinary plasticity to adapt to different
agroecosystems. For example, Palmer amaranth mimicked crop architecture
in competing for light (Figure 6), where plants growing in corn
allocated resources to height, while plants growing in bareground
allocated resources to numerous branches. Plants growing in the absence
of a crop produced the greatest biomass, responding to a greater
abundance of light, nutrient and water resources (Figure 3A, 3B). These
results support the argument that Palmer amaranth can quickly evolve
life-history traits to adapt to different cultural practices, similar to
observations in a Palmer amaranth response to nitrogen study (Bravo et
al., 2018). Our results highlight Palmer amaranth as a threat to field
crops as breeding more competitive crop varieties is likely to select
more competitive weed biotypes (Bravo et al., 2017).

Palmer amaranth growth and development in second cohort was limited due
to the crop competitive ability at advanced development stages. Plants
were transplanted at greater crop height and width, which reduced Palmer
amaranth competitiveness. As a result, Palmer amaranth height and
biomass was lower compared to its first cohort. Moreover, Palmer
amaranth growing without crop competition produced the highest amounts
of biomass. The Palmer amaranth strategy in bareground was to invest
biomass in growing plant width and height. Nonetheless, Palmer amaranth
produced 17\% less biomass in second cohort compared to first cohort
timing. In a bareground study, early emerged Palmer amaranth without
competition was 50\% taller than late emerged plants (Webster and Grey,
2015). These results suggest that crop competition is not the only
factor limiting late Palmer amaranth establishment. The limited growth
of Palmer amaranth at second cohort is likely a reduced plant response
to day length, light availability and thermal units (e.g, growing degree
days). The \emph{Amaranthus} species are sensitive to photoperiod (Wu
and Owen, 2014). It is hypothesize that reduced day length or
red/far-red effect contributed to smaller plants at second cohort
regardless the crop. A study in North Carolina and Illinois predicted
that less than 10\% Palmer amaranth seedlings emergence occurred after
June (Piskackova et al., 2021). In addition, Palmer amaranth negative
impact on soybean (Korres et al., 2020) and cotton (Webster and Grey,
2015) yields was higher when plants were established near to crop
planting. Therefore, early management is a key strategy to minimize the
damaging impact of Palmer amaranth to US Midwest cropping systems.

Seed production was not evaluated due to plant harvest at initiation of
flowering. Nonetheless, it is well documented a strong positive
correlation between Palmer amaranth biomass and seed production
(Schwartz et al., 2016; Spaunhorst et al., 2018). In our study, plants
growing from the first cohort accumulated 42\% more biomass when
compared to the second cohort. Therefore, Palmer amaranth plants growing
in the second cohort is likely to produce less seeds regardless the
crop. Our observation is consistent with the findings that the first
Palmer amaranth cohort produced 50\% more seeds per plant than Palmer
amaranth plants established six weeks later in bareground (Webster and
Grey, 2015). Still, seed production at the second cohort will replenish
the soil seedbank. Seed production and deposition in the seedbank is
also a key factor for species perpetuation (Menges, 1987). Palmer
amaranth can produce a hundred thousands seeds per plant (Schwartz et
al., 2016; Keeley et al., 1987), which can stay viable in the soil
seedbank for at least 36 months (Sosnoskie et al., 2013). Therefore,
preventing Palmer amaranth seed production or/and seed migration to
non-native habitats is an essential strategy to minimize the species
impact in agroecosystems (Davis et al., 2015).

Floral initiation depends on the complex interaction between a plant's
genetic makeup and environmental conditions (Lang, 1965). We observed a
significant disruption in flowering caused by surrounding vegetation and
cohort timing. Although all three treatments began flowering at the same
time in the first cohort (about doy 180), it took much less time for the
bareground treatment to reach 90\% flowering (34 days versus 40 for
soybean and 71 for corn). The lack of competition (and stress) in the
bareground resulted in plants that flowered earlier, and were they not
harvested, would have had longer reproductive periods, thereby producing
more seed. Similarly, in the second cohort, the window for all plants to
initiate flowering was much shorter for the bareground (28 days)
compared to Palmer amaranth in soybean (50 days) or corn (44 days).
Significantly, plants in the bareground treatment flowered one week
earlier than soybean, and almost two weeks earlier than corn, again
allowing for greater seed production per plant had they not been
harvested at initial flowering. In some cases, early flowering may
confer an evolutionary advantage, provided the plant has an
indeterminate habit and flowering does not restrict plant growth. In an
Iowa study, Palmer amaranth initiated flowering two weeks prior to
waterhemp (Baker, 2021), and this characteristic may partially explain
why Palmer amaranth can displace waterhemp where both occur in the same
agronomic field. In other cases, however, early flower initiation may be
a response to stressful conditions as a plant attempts to reproduce
before running out of resources. In a dioecious species like Palmer
amaranth, exerting stress on plants to manipulate flowering may be
beneficial in limiting seed production (McFarlane et al., 2018;
Schliekelman et al., 2005). For example, when growing under water
stress, there was a seven days flowering mismatch between male and
female plants (Mesgaran et al., 2021), which can minimize plant
outcrossing, reduce seed production and the exchange of resistant
alleles (Jhala et al., 2021). In our study we observed

An ecological approach to reduce Palmer amaranth seed production is
understanding plant biology, including flowering pattern. Our study
suggests that Palmer amaranth flowering was slightly influenced by crops
and cohort timings. Palmer amaranth growing in bareground and corn
resulted in the overall shortest and longest flowering window,
respectively. The shortest flowering window of second cohorts of Palmer
amaranth growing in bareground highlighted the impact of photoperiod on
flowering. When growing in soybean, Palmer amaranth flowering window was
similar to bareground for the first cohort but similar to corn at second
cohort timing. Plant flowering initiation is complex and depends on the
complex interaction between genetic makeup and the environmental
conditions (Lang, 1965). We hypothesize that when growing in high
competition (e.g., second cohort), Palmer amaranth plants tend to
initiate flowering early, as well as having an extended flowering
window. A study has shown that Palmer amaranth initiated flowering two
weeks prior to the native waterhemp in Iowa (Baker, 2021). Early flower
initiation could also be a plant strategy when growing in stressful
conditions. For example, when growing under water stress, early
flowering in Palmer amaranth resulted in a mismatch between female and
male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
amaranth male and female flowering period can minimize plant
outcrossing, reducing plant seed production and/or exchange of resistant
alleles (Jhala et al., 2021). Sex dimorphism manipulation is considered
a potential ecological pest control strategy (McFarlane et al., 2018;
Schliekelman et al., 2005).

The mechanisms of gender-determination in plant species is intriguing
and has arouse the curiosity of many scientists, including Charles
Darwin (Darwin, 1888). In our study, the gender model performance was
decent (AUC 0.64) considering the biology of plant flowering. A 1:1 male
and female sex ratio is a general an evolutionarily stable strategy for
plant species perpetuation (Fisher, 1930). However, a slight deviation
from 1:1 sex ratio occurs in some dioecious species. For example, the
dioecious \emph{Halophila stipulacea} is a female-biased plant in its
native habitat, but the naturalized \emph{H. stipulacea} has a 1:1 ratio
(Nguyen et al., 2018). Naturalization of \emph{H. stipulacea} reduced
female-male ratio to expand into its non-native habitat (Nguyen et al.,
2018). Also, biotic and/or abiotic stress can influence plant gender
determination. Palmer amaranth male-to-female ratio was greater under
high plant densities (Korres and Norsworthy, 2017) and after herbicide
application (Rumpa et al., 2019). We observed sexual dimorphism in
Palmer amaranth in response to surrounding vegetation and plant
morphological attributes. Our model estimated that late flowering,
heavier and taller Palmer amaranth plants slightly deviated from 1:1
ratio in favor of female plants. It was reported that female Palmer
amaranth plants invested more in height, stem and biomass while male
plants invested more in leaf area and leaf dry weight under nutrient
deficiency (Korres et al., 2017). We observed more female plants in
soybean and bareground compared to corn. Palmer amaranth plants in the
corn were more stressed by interspecific competition as evidenced by
less biomass and a lower weight:height ratio. Sexual dimorphism is
documented in other dioecious species (Barrett and Hough, 2013). For
example, stronger female plant competition and greater male tolerance to
herbivory was reported in \emph{Spinacia oleracea} (Pérez-Llorca and
Sánchez Vilas, 2019). Research on candidate genes for sex determination
in \emph{Amaranthus} species are currently underway but is far from
complete (Montgomery et al., 2021, 2019). Further studies are also
needed to understand the ecological basis of Palmer amaranth flowering,
including plant behavior under climate change.

Our study demonstrated that Palmer amaranth is adapted to grow on arable
land throughout the Midwestern U.S. Palmer amaranth's range will
continue to expand if current cropping practices are continued. As
waterhemp and Palmer amaranth begin to share the same habitat, it will
increase weed management complexity. Preventing Palmer amaranth seed
dispersal must be a priority. Regional collaboration is necessary to
slow the spread of this aggressive and adaptable weed. Where Palmer
amaranth occurs, management tactics should focus on limiting Palmer
amaranth establishment until row crops can shade late-emerging plants.
Increasing the diversity of crops in rotation, varying row crop planting
date, narrowing row width, and increasing residue cover through the use
of cover crops are all tactics that can minimize the growth and seed
production of Palmer amaranth, and improve the sustainability of
Midwestern cropping systems.

\hypertarget{disclosureconflict-of-interest-statement}{%
\section*{Disclosure/Conflict-of-Interest
Statement}\label{disclosureconflict-of-interest-statement}}
\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\hypertarget{data-accessibility}{%
\section*{Data accessibility}\label{data-accessibility}}
\addcontentsline{toc}{section}{Data accessibility}

The data and scripts used to analyse the data presented in this work can
be found at Zenodo (Oliveira, 2021).

\hypertarget{author-contributions}{%
\section*{Author Contributions}\label{author-contributions}}
\addcontentsline{toc}{section}{Author Contributions}

RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted
the experiments; MO: analyzed the data; MO: wrote the manuscript and all
authors revised; AJ, CP, MB, MO, SS, and RW: conceptualized the
research. All authors reviewed the manuscript.

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

Funding: This work received no specific grant from any funding agency,
commercial, or not-for-profit sectors.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-bagavathiannan2016}{}%
Bagavathiannan, M. V., and Norsworthy, J. K. (2016). Multiple-{Herbicide
Resistance Is Widespread} in {Roadside Palmer Amaranth Populations}.
\emph{PLOS ONE} 11, e0148748.
doi:\href{https://doi.org/10.1371/journal.pone.0148748}{10.1371/journal.pone.0148748}.

\leavevmode\hypertarget{ref-baker2021}{}%
Baker, R. (2021). Comparative analysis of {Palmer} amaranth
({Amaranthus} palmeri) and waterhemp ({A}. Tuberculatus) in {Iowa}.
doi:\href{https://doi.org/10.31274/etd-20210609-11}{10.31274/etd-20210609-11}.

\leavevmode\hypertarget{ref-bangdiwala2018}{}%
Bangdiwala, S. I. (2018). Regression: Binary logistic.
\emph{International Journal of Injury Control and Safety Promotion} 25,
336--338.
doi:\href{https://doi.org/10.1080/17457300.2018.1486503}{10.1080/17457300.2018.1486503}.

\leavevmode\hypertarget{ref-barrett2013}{}%
Barrett, S. C. H., and Hough, J. (2013). Sexual dimorphism in flowering
plants. \emph{Journal of Experimental Botany} 64, 67--82.
doi:\href{https://doi.org/10.1093/jxb/ers308}{10.1093/jxb/ers308}.

\leavevmode\hypertarget{ref-bates2015}{}%
Bates, D., Mächler, M., Bolker, B., and Walker, S. (2015). Fitting
{Linear Mixed}-{Effects Models Using} Lme4. \emph{Journal of Statistical
Software} 67, 1--48.
doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}.

\leavevmode\hypertarget{ref-berger2015}{}%
Berger, S. T., Ferrell, J. A., Rowland, D. L., and Webster, T. M.
(2015). Palmer {Amaranth} ({Amaranthus} palmeri) {Competition} for
{Water} in {Cotton}. \emph{Weed Science} 63, 928--935.
doi:\href{https://doi.org/10.1614/WS-D-15-00062.1}{10.1614/WS-D-15-00062.1}.

\leavevmode\hypertarget{ref-bravo2017}{}%
Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
(2017). Differentiation of {Life}-{History Traits} among {Palmer
Amaranth Populations} ({Amaranthus} palmeri) and {Its Relation} to
{Cropping Systems} and {Glyphosate Sensitivity}. \emph{Weed Science} 65,
339--349.
doi:\href{https://doi.org/10.1017/wsc.2017.14}{10.1017/wsc.2017.14}.

\leavevmode\hypertarget{ref-bravo2018}{}%
Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
(2018). Evolutionary {Adaptations} of {Palmer Amaranth} ({Amaranthus}
palmeri) to {Nitrogen Fertilization} and {Crop Rotation History Affect
Morphology} and {Nutrient}-{Use Efficiency}. \emph{Weed Science} 66,
180--189.
doi:\href{https://doi.org/10.1017/wsc.2017.73}{10.1017/wsc.2017.73}.

\leavevmode\hypertarget{ref-briscoerunquist2019}{}%
Briscoe Runquist, R. D., Lake, T., Tiffin, P., and Moeller, D. A.
(2019). Species distribution models throughout the invasion history of
{Palmer} amaranth predict regions at risk of future invasion and reveal
challenges with modeling rapidly shifting geographic ranges. \emph{Sci
Rep} 9, 2426.
doi:\href{https://doi.org/10.1038/s41598-018-38054-9}{10.1038/s41598-018-38054-9}.

\leavevmode\hypertarget{ref-chahal2021}{}%
Chahal, P. S., Barnes, E. R., and Jhala, A. J. (2021). Emergence pattern
of {Palmer} amaranth ({Amaranthus} palmeri) influenced by tillage
timings and residual herbicides. \emph{Weed Technology} 35, 433--439.
doi:\href{https://doi.org/10.1017/wet.2020.136}{10.1017/wet.2020.136}.

\leavevmode\hypertarget{ref-chahal2018}{}%
Chahal, P. S., Irmak, S., Jugulam, M., and Jhala, A. J. (2018).
Evaluating {Effect} of {Degree} of {Water Stress} on {Growth} and
{Fecundity} of {Palmer} amaranth ({Amaranthus} palmeri) {Using Soil
Moisture Sensors}. \emph{Weed Science} 66, 738--745.
doi:\href{https://doi.org/10.1017/wsc.2018.47}{10.1017/wsc.2018.47}.

\leavevmode\hypertarget{ref-correndo2021}{}%
Correndo, A. A., Moro Rosso, L. H., and Ciampitti, I. A. (2021).
Retrieving and processing agro-meteorological data from {API}-client
sources using {R} software. \emph{BMC Research Notes} 14, 205.
doi:\href{https://doi.org/10.1186/s13104-021-05622-8}{10.1186/s13104-021-05622-8}.

\leavevmode\hypertarget{ref-crespo2016a}{}%
Crespo, R. J., Wingeyer, A. B., Borman, C. J., and Bernards, M. L.
(2016). Baseline {Sensitivity} of {Nebraska Waterhemp} and {Palmer
Amaranth} to {Dicamba} and 2,4-{D}. \emph{Agronomy Journal} 108,
1649--1655.
doi:\href{https://doi.org/10.2134/agronj2015.0465}{10.2134/agronj2015.0465}.

\leavevmode\hypertarget{ref-darwin1888}{}%
Darwin, C. (1888). \emph{The {Different Forms} of {Flowers} on {Plants}
of the {Same Species}}. {J. Murray} Available at:
\url{http://books.google.com?id=7uMEAAAAYAAJ}.

\leavevmode\hypertarget{ref-davis2015}{}%
Davis, A. S., Schutte, B. J., Hager, A. G., and Young, B. G. (2015).
Palmer {Amaranth} ({Amaranthus} palmeri) {Damage Niche} in {Illinois
Soybean Is Seed Limited}. \emph{Weed Science} 63, 658--668.
doi:\href{https://doi.org/10.1614/WS-D-14-00177.1}{10.1614/WS-D-14-00177.1}.

\leavevmode\hypertarget{ref-farmer2017}{}%
Farmer, J. A., Webb, E. B., Pierce, R. A., and Bradley, K. W. (2017).
Evaluating the potential for weed seed dispersal based on waterfowl
consumption and seed viability. \emph{Pest Management Science} 73,
2592--2603. doi:\href{https://doi.org/10.1002/ps.4710}{10.1002/ps.4710}.

\leavevmode\hypertarget{ref-fisher1930}{}%
Fisher, R. A. (1930). The genetical theory of natural selection.
\emph{Eugen Rev} 22, 127--130. Available at:
\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2984947/} {[}Accessed
August 12, 2021{]}.

\leavevmode\hypertarget{ref-fox2018}{}%
Fox, J., and Weisberg, S. (2018). \emph{An {R Companion} to {Applied
Regression}}. {SAGE Publications} Available at:
\url{http://books.google.com?id=uPNrDwAAQBAJ}.

\leavevmode\hypertarget{ref-garetson2019}{}%
Garetson, R., Singh, V., Singh, S., Dotray, P., and Bagavathiannan, M.
(2019). Distribution of herbicide-resistant {Palmer} amaranth
({Amaranthus} palmeri) in row crop production systems in {Texas}.
\emph{Weed Technology} 33, 355--365.
doi:\href{https://doi.org/10.1017/wet.2019.14}{10.1017/wet.2019.14}.

\leavevmode\hypertarget{ref-guo2003}{}%
Guo, P., and Al-Khatib, K. (2003). Temperature effects on germination
and growth of redroot pigweed ({Amaranthus} retroflexus), {Palmer}
amaranth ({A}. Palmeri), and common waterhemp ({A}. rudis). \emph{Weed
Science} 51, 869--875.
doi:\href{https://doi.org/10.1614/P2002-127}{10.1614/P2002-127}.

\leavevmode\hypertarget{ref-hartzler2016}{}%
Hartzler, B., and Anderson, M. (2016). Palmer amaranth: {It}'s here, now
what? 10.

\leavevmode\hypertarget{ref-heap2021}{}%
Heap, I. (2021). Internation {Herbicide}-{Resistant Weed Database}.
Available at: \url{http://www.weedscience.org/Home.aspx} {[}Accessed
July 26, 2021{]}.

\leavevmode\hypertarget{ref-hothorn2008}{}%
Hothorn, T., Bretz, F., and Westfall, P. (2008). Simultaneous
{Inference} in {General Parametric Models}. \emph{Biometrical Journal}
50, 346--363.
doi:\href{https://doi.org/10.1002/bimj.200810425}{10.1002/bimj.200810425}.

\leavevmode\hypertarget{ref-jha2010}{}%
Jha, P., Norsworthy, J. K., Riley, M. B., and Bridges, W. (2010). Annual
{Changes} in {Temperature} and {Light Requirements} for {Germination} of
{Palmer Amaranth} ({Amaranthus} palmeri) {Seeds Retrieved} from {Soil}.
\emph{Weed Science} 58, 426--432.
doi:\href{https://doi.org/10.1614/WS-D-09-00038.1}{10.1614/WS-D-09-00038.1}.

\leavevmode\hypertarget{ref-jhala2021}{}%
Jhala, A. J., Norsworthy, J. K., Ganie, Z. A., Sosnoskie, L. M., Beckie,
H. J., Mallory-Smith, C. A., Liu, J., Wei, W., Wang, J., and
Stoltenberg, D. E. (2021). Pollen-mediated gene flow and transfer of
resistance alleles from herbicide-resistant broadleaf weeds. \emph{Weed
Technology} 35, 173--187.
doi:\href{https://doi.org/10.1017/wet.2020.101}{10.1017/wet.2020.101}.

\leavevmode\hypertarget{ref-keeley1987}{}%
Keeley, P. E., Carter, C. H., and Thullen, R. J. (1987). Influence of
{Planting Date} on {Growth} of {Palmer Amaranth} ({Amaranthus} palmeri).
\emph{Weed Science} 35, 199--204.
doi:\href{https://doi.org/10.1017/S0043174500079054}{10.1017/S0043174500079054}.

\leavevmode\hypertarget{ref-kistner2018}{}%
Kistner, E. J., and Hatfield, J. L. (2018). Potential {Geographic
Distribution} of {Palmer Amaranth} under {Current} and {Future
Climates}. \emph{Agricultural \& Environmental Letters} 3, 170044.
doi:\href{https://doi.org/10.2134/ael2017.12.0044}{10.2134/ael2017.12.0044}.

\leavevmode\hypertarget{ref-klingaman1994}{}%
Klingaman, T. E., and Oliver, L. R. (1994). Palmer {Amaranth}
({Amaranthus} palmeri) {Interference} in {Soybeans} ({Glycine} max).
\emph{Weed Science} 42, 523--527.
doi:\href{https://doi.org/10.1017/S0043174500076888}{10.1017/S0043174500076888}.

\leavevmode\hypertarget{ref-korres2017a}{}%
Korres, N. E., and Norsworthy, J. K. (2017). Palmer {Amaranth}
({Amaranthus} palmeri) {Demographic} and {Biological Characteristics} in
{Wide}-{Row Soybean}. \emph{Weed Science} 65, 491--503.
doi:\href{https://doi.org/10.1017/wsc.2017.12}{10.1017/wsc.2017.12}.

\leavevmode\hypertarget{ref-korres2017}{}%
Korres, N. E., Norsworthy, J. K., FitzSimons, T., Roberts, T. L., and
Oosterhuis, D. M. (2017). Differential {Response} of {Palmer Amaranth}
({Amaranthus} palmeri) {Gender} to {Abiotic Stress}. \emph{Weed Science}
65, 213--227.
doi:\href{https://doi.org/10.1017/wsc.2016.34}{10.1017/wsc.2016.34}.

\leavevmode\hypertarget{ref-korres2020}{}%
Korres, N. E., Norsworthy, J. K., Mauromoustakos, A., and Williams, M.
M. (2020). Soybean density and {Palmer} amaranth ({Amaranthus} palmeri)
establishment time: Effects on weed biology, crop yield, and economic
returns. \emph{Weed Science} 68, 467--475.
doi:\href{https://doi.org/10.1017/wsc.2020.41}{10.1017/wsc.2020.41}.

\leavevmode\hypertarget{ref-kuhn2021}{}%
Kuhn, M., Vaughan, D., and RStudio (2021). \emph{Yardstick: {Tidy
Characterizations} of {Model Performance}}. Available at:
\url{https://CRAN.R-project.org/package=yardstick} {[}Accessed August
24, 2021{]}.

\leavevmode\hypertarget{ref-kumar2020}{}%
Kumar, V., Liu, R., and Stahlman, P. W. (2020). Differential sensitivity
of {Kansas Palmer} amaranth populations to multiple herbicides.
\emph{Agronomy Journal} 112, 2152--2163.
doi:\href{https://doi.org/10.1002/agj2.20178}{10.1002/agj2.20178}.

\leavevmode\hypertarget{ref-kupper2017}{}%
Küpper, A., Borgato, E. A., Patterson, E. L., Netto, A. G., Nicolai, M.,
Carvalho, S. J. P. de, Nissen, S. J., Gaines, T. A., and Christoffoleti,
P. J. (2017). Multiple {Resistance} to {Glyphosate} and {Acetolactate
Synthase Inhibitors} in {Palmer Amaranth} ({Amaranthus} palmeri)
{Identified} in {Brazil}. \emph{Weed Science} 65, 317--326.
doi:\href{https://doi.org/10.1017/wsc.2017.1}{10.1017/wsc.2017.1}.

\leavevmode\hypertarget{ref-lang1965}{}%
Lang, A. (1965). {``Physiology of flower initiation,''} in
\emph{Differenzierung und {Entwicklung} / {Differentiation} and
{Development}} Handbuch der {Pflanzenphysiologie} / {Encyclopedia} of
{Plant Physiology}., ed. A. Lang ({Berlin, Heidelberg}: {Springer}),
1380--1536.
doi:\href{https://doi.org/10.1007/978-3-642-50088-6_39}{10.1007/978-3-642-50088-6\_39}.

\leavevmode\hypertarget{ref-larran2017}{}%
Larran, A. S., Palmieri, V. E., Perotti, V. E., Lieber, L., Tuesca, D.,
and Permingeat, H. R. (2017). Target-site resistance to acetolactate
synthase ({ALS})-inhibiting herbicides in {Amaranthus} palmeri from
{Argentina}. \emph{Pest Management Science} 73, 2578--2584.
doi:\href{https://doi.org/10.1002/ps.4662}{10.1002/ps.4662}.

\leavevmode\hypertarget{ref-leeper2017}{}%
Leeper, T. J. (2017). Interpreting {Regression Results} using {Average
Marginal Eﬀects} with {R}'s margins. 31.

\leavevmode\hypertarget{ref-leeper2021}{}%
Leeper, T. J., Arnold, J., Arel-Bundock, V., and Long, J. A. (2021).
\emph{Margins: {Marginal Effects} for {Model Objects}}. Available at:
\url{https://CRAN.R-project.org/package=margins} {[}Accessed August 24,
2021{]}.

\leavevmode\hypertarget{ref-lenth2021}{}%
Lenth, R. V., Buerkner, P., Herve, M., Love, J., Riebl, H., and
Singmann, H. (2021). \emph{Emmeans: {Estimated Marginal Means}, aka
{Least}-{Squares Means}}. Available at:
\url{https://CRAN.R-project.org/package=emmeans} {[}Accessed August 24,
2021{]}.

\leavevmode\hypertarget{ref-lindsay2017}{}%
Lindsay, K., Popp, M., Norsworthy, J., Bagavathiannan, M., Powles, S.,
and Lacoste, M. (2017). {PAM}: {Decision Support} for {Long}-{Term
Palmer Amaranth} ({Amaranthus} palmeri) {Control}. \emph{Weed
Technology} 31, 915--927.
doi:\href{https://doi.org/10.1017/wet.2017.69}{10.1017/wet.2017.69}.

\leavevmode\hypertarget{ref-macrae2013}{}%
MacRae, A. W., Webster, T. M., Sosnoskie, L. M., Culpepper, A. S., and
Kichler, J. M. (2013). Cotton {Yield Loss Potential} in {Response} to
{Length} of {Palmer Amaranth} ({Amaranthus} palmeri) {Interference}. 17,
6.

\leavevmode\hypertarget{ref-mangiafico2021}{}%
Mangiafico, S. (2021). \emph{Rcompanion: {Functions} to {Support
Extension Education Program Evaluation}}. Available at:
\url{https://CRAN.R-project.org/package=rcompanion} {[}Accessed August
24, 2021{]}.

\leavevmode\hypertarget{ref-massinga2001}{}%
Massinga, R. A., Currie, R. S., Horak, M. J., and Boyer, J. (2001).
Interference of {Palmer} amaranth in corn. \emph{Weed Science} 49,
202--208.
doi:\href{https://doi.org/10.1614/0043-1745(2001)049\%5B0202:IOPAIC\%5D2.0.CO;2}{10.1614/0043-1745(2001)049{[}0202:IOPAIC{]}2.0.CO;2}.

\leavevmode\hypertarget{ref-matzrafi2021}{}%
Matzrafi, M., Osipitan, O. A., Ohadi, S., and Mesgaran, M. B. (2021).
Under pressure: Maternal effects promote drought tolerance in progeny
seed of {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed Science}
69, 31--38.
doi:\href{https://doi.org/10.1017/wsc.2020.75}{10.1017/wsc.2020.75}.

\leavevmode\hypertarget{ref-mcfarlane2018}{}%
McFarlane, G. R., Whitelaw, C. B. A., and Lillico, S. G. (2018).
{CRISPR}-{Based Gene Drives} for {Pest Control}. \emph{Trends in
Biotechnology} 36, 130--133.
doi:\href{https://doi.org/10.1016/j.tibtech.2017.10.001}{10.1016/j.tibtech.2017.10.001}.

\leavevmode\hypertarget{ref-menges1987}{}%
Menges, R. M. (1987). Weed {Seed Population Dynamics} during {Six Years}
of {Weed Management Systems} in {Crop Rotations} on {Irrigated Soil}.
\emph{Weed Science} 35, 328--332. Available at:
\url{http://www.jstor.org/stable/4044593}.

\leavevmode\hypertarget{ref-mesgaran2021}{}%
Mesgaran, M. B., Matzrafi, M., and Ohadi, S. (2021). Sex dimorphism in
dioecious {Palmer} amaranth ({Amaranthus} palmeri) in response to water
stress. \emph{Planta} 254, 17.
doi:\href{https://doi.org/10.1007/s00425-021-03664-7}{10.1007/s00425-021-03664-7}.

\leavevmode\hypertarget{ref-milani2021}{}%
Milani, A., Panozzo, S., Farinati, S., Iamonico, D., Sattin, M., Loddo,
D., and Scarabel, L. (2021). Recent {Discovery} of {Amaranthus} palmeri
{S}. {Watson} in {Italy}: {Characterization} of {ALS}-{Resistant
Populations} and {Sensitivity} to {Alternative Herbicides}.
\emph{Sustainability} 13, 7003.
doi:\href{https://doi.org/10.3390/su13137003}{10.3390/su13137003}.

\leavevmode\hypertarget{ref-montgomery2021}{}%
Montgomery, J. S., Giacomini, D. A., Weigel, D., and Tranel, P. J.
(2021). Male-specific {Y}-chromosomal regions in waterhemp ({Amaranthus}
tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{New
Phytologist} 229, 3522--3533.
doi:\href{https://doi.org/10.1111/nph.17108}{10.1111/nph.17108}.

\leavevmode\hypertarget{ref-montgomery2019}{}%
Montgomery, J. S., Sadeque, A., Giacomini, D. A., Brown, P. J., and
Tranel, P. J. (2019). Sex-specific markers for waterhemp ({Amaranthus}
tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed
Science} 67, 412--418.
doi:\href{https://doi.org/10.1017/wsc.2019.27}{10.1017/wsc.2019.27}.

\leavevmode\hypertarget{ref-morgan2001}{}%
Morgan, G. D., Baumann, P. A., and Chandler, J. M. (2001). Competitive
{Impact} of {Palmer Amaranth} ({Amaranthus} palmeri) on {Cotton}
({Gossypium} hirsutum) {Development} and {Yield}. \emph{Weed Technology}
15, 408--412.
doi:\href{https://doi.org/10.1614/0890-037X(2001)015\%5B0408:CIOPAA\%5D2.0.CO;2}{10.1614/0890-037X(2001)015{[}0408:CIOPAA{]}2.0.CO;2}.

\leavevmode\hypertarget{ref-nguyen2018}{}%
Nguyen, H. M., Kleitou, P., Kletou, D., Sapir, Y., and Winters, G.
(2018). Differences in flowering sex ratios between native and invasive
populations of the seagrass {Halophila} stipulacea. \emph{Botanica
Marina} 61, 337--342.
doi:\href{https://doi.org/10.1515/bot-2018-0015}{10.1515/bot-2018-0015}.

\leavevmode\hypertarget{ref-oliveira2021b}{}%
Oliveira, M. C. (2021). \emph{Maxwelco/palmer\_adaptation}. {Zenodo}
Available at: \url{https://doi.org/10.5281/zenodo.5236831}.

\leavevmode\hypertarget{ref-oliveira2018}{}%
Oliveira, M. C., Gaines, T. A., Patterson, E. L., Jhala, A. J., Irmak,
S., Amundsen, K., and Knezevic, S. Z. (2018). Interspecific and
intraspecific transference of metabolism-based mesotrione resistance in
dioecious weedy {Amaranthus}. \emph{The Plant Journal} 96, 1051--1063.
doi:\href{https://doi.org/10.1111/tpj.14089}{10.1111/tpj.14089}.

\leavevmode\hypertarget{ref-oliveira2021a}{}%
Oliveira, M. C., Giacomini, D. A., Arsenijevic, N., Vieira, G., Tranel,
P. J., and Werle, R. (2021). Distribution and validation of genotypic
and phenotypic glyphosate and {PPO}-inhibitor resistance in {Palmer}
amaranth ({Amaranthus} palmeri) from southwestern {Nebraska}. \emph{Weed
Technology} 35, 65--76.
doi:\href{https://doi.org/10.1017/wet.2020.74}{10.1017/wet.2020.74}.

\leavevmode\hypertarget{ref-page2021}{}%
Page, E. R., Nurse, R. E., Meloche, S., Bosveld, K., Grainger, C.,
Obeid, K., Filotas, M., Simard, M.-J., and Laforest, M. (2021). Import
of {Palmer} amaranth ({Amaranthus} palmeri {S}. {Wats}.) Seed with sweet
potato ({Ipomea} batatas ({L}.) {Lam}) slips. \emph{Can. J. Plant Sci.},
CJPS-2020-0321.
doi:\href{https://doi.org/10.1139/CJPS-2020-0321}{10.1139/CJPS-2020-0321}.

\leavevmode\hypertarget{ref-perez-llorca2019}{}%
Pérez-Llorca, M., and Sánchez Vilas, J. (2019). Sexual dimorphism in
response to herbivory and competition in the dioecious herb {Spinacia}
oleracea. \emph{Plant Ecol} 220, 57--68.
doi:\href{https://doi.org/10.1007/s11258-018-0902-7}{10.1007/s11258-018-0902-7}.

\leavevmode\hypertarget{ref-piskackova2021}{}%
Piskackova, T. A. R., Reberg-Horton, S. C., Richardson, R. J., Jennings,
K. M., Franca, L., Young, B. G., and Leon, R. G. (2021). Windows of
action for controlling palmer amaranth ({Amaranthus} palmeri) using
emergence and phenology models. \emph{Weed Research} 61, 188--198.
doi:\href{https://doi.org/10.1111/wre.12470}{10.1111/wre.12470}.

\leavevmode\hypertarget{ref-price2011}{}%
Price, A. J., Balkcom, K. S., Culpepper, S. A., Kelton, J. A., Nichols,
R. L., and Schomberg, H. (2011). Glyphosate-resistant {Palmer} amaranth:
{A} threat to conservation tillage. \emph{Journal of Soil and Water
Conservation} 66, 265--275.
doi:\href{https://doi.org/10.2489/jswc.66.4.265}{10.2489/jswc.66.4.265}.

\leavevmode\hypertarget{ref-raoniar2021}{}%
Raoniar, R. (2021). Modelling {Binary Logistic Regression Using R}
(research-oriented modelling and interpretation). Available at:
\url{https://bit.ly/3BhucN3} {[}Accessed August 11, 2021{]}.

\leavevmode\hypertarget{ref-ritz2015}{}%
Ritz, C., Baty, F., Streibig, J. C., and Gerhard, D. (2015).
Dose-{Response Analysis Using R}. \emph{PLOS ONE} 10, e0146021.
doi:\href{https://doi.org/10.1371/journal.pone.0146021}{10.1371/journal.pone.0146021}.

\leavevmode\hypertarget{ref-rumpa2019}{}%
Rumpa, M. M., Krausz, R. F., Gibson, D. J., and Gage, K. L. (2019).
Effect of {PPO}-{Inhibiting Herbicides} on the {Growth} and {Sex Ratio}
of a {Dioecious Weed Species Amaranthus} palmeri ({Palmer Amaranth}).
\emph{Agronomy} 9, 275.
doi:\href{https://doi.org/10.3390/agronomy9060275}{10.3390/agronomy9060275}.

\leavevmode\hypertarget{ref-sauer1957}{}%
Sauer, J. (1957). Recent {Migration} and {Evolution} of the {Dioecious
Amaranths}. \emph{Evolution} 11, 11--31.
doi:\href{https://doi.org/10.2307/2405808}{10.2307/2405808}.

\leavevmode\hypertarget{ref-sauer1972}{}%
Sauer, J. D. (1972). The dioecious amaranths: A new species name and
major range extensions. \emph{Madroño} 21, 426--434. Available at:
\url{http://www.jstor.org/stable/41423815}.

\leavevmode\hypertarget{ref-sbatella2010}{}%
Sbatella, G. M., and Wilson, R. G. (2010). Isoxaflutole {Shifts Kochia}
({Kochia} scoparia) {Populations} in {Continuous Corn}. \emph{Weed
Technology} 24, 392--396.
doi:\href{https://doi.org/10.1614/WT-D-09-00023.1}{10.1614/WT-D-09-00023.1}.

\leavevmode\hypertarget{ref-schliekelman2005}{}%
Schliekelman, P., Ellner, S., and Gould, F. (2005). Pest {Control} by
{Genetic Manipulation} of {Sex Ratio}. \emph{Journal of Economic
Entomology} 98, 18--34.
doi:\href{https://doi.org/10.1093/jee/98.1.18}{10.1093/jee/98.1.18}.

\leavevmode\hypertarget{ref-schwartz2016}{}%
Schwartz, L. M., Norsworthy, J. K., Young, B. G., Bradley, K. W.,
Kruger, G. R., Davis, V. M., Steckel, L. E., and Walsh, M. J. (2016).
Tall {Waterhemp} ({Amaranthus} tuberculatus) and {Palmer} amaranth
({Amaranthus} palmeri) {Seed Production} and {Retention} at {Soybean
Maturity}. \emph{Weed Technology} 30, 284--290.
doi:\href{https://doi.org/10.1614/WT-D-15-00130.1}{10.1614/WT-D-15-00130.1}.

\leavevmode\hypertarget{ref-sing2005}{}%
Sing, T., Sander, O., Beerenwinkel, N., and Lengauer, T. (2005). {ROCR}:
Visualizing classifier performance in {R}. \emph{Bioinformatics} 21,
3940--3941.
doi:\href{https://doi.org/10.1093/bioinformatics/bti623}{10.1093/bioinformatics/bti623}.

\leavevmode\hypertarget{ref-sosnoskie2013}{}%
Sosnoskie, L. M., Webster, T. M., and Culpepper, A. S. (2013).
Glyphosate {Resistance Does Not Affect Palmer Amaranth} ({Amaranthus}
palmeri) {Seedbank Longevity}. \emph{Weed Science} 61, 283--288.
doi:\href{https://doi.org/10.1614/WS-D-12-00111.1}{10.1614/WS-D-12-00111.1}.

\leavevmode\hypertarget{ref-spaunhorst2018}{}%
Spaunhorst, D. J., Devkota, P., Johnson, W. G., Smeda, R. J., Meyer, C.
J., and Norsworthy, J. K. (2018). Phenology of {Five Palmer} amaranth
({Amaranthus} palmeri) {Populations Grown} in {Northern Indiana} and
{Arkansas}. \emph{Weed Science} 66, 457--469.
doi:\href{https://doi.org/10.1017/wsc.2018.12}{10.1017/wsc.2018.12}.

\leavevmode\hypertarget{ref-rcoreteam2021}{}%
Team, R. C. (2021). \emph{R: {The R Project} for {Statistical
Computing}}. {Vienna, Austria}: {R Foundation for Statistical Computing}
Available at: \url{https://www.r-project.org/} {[}Accessed August 24,
2021{]}.

\leavevmode\hypertarget{ref-vanwychen2020}{}%
Van Wychen, L. (2020). 2020 {Survey} of the most common and troublesome
weeds in grass crops, pasture, and turf in the {United States} and
{Canada}. Available at:
\url{https://wssa.net/wp-content/uploads/2020-Weed-Survey_grass-crops.xlsx}.

\leavevmode\hypertarget{ref-wang1992}{}%
Wang, J. L., Klessig, D. F., and Berry, J. O. (1992). Regulation of {C4
Gene Expression} in {Developing Amaranth Leaves}. \emph{The Plant Cell}
4, 173--184.
doi:\href{https://doi.org/10.1105/tpc.4.2.173}{10.1105/tpc.4.2.173}.

\leavevmode\hypertarget{ref-ward2013}{}%
Ward, S. M., Webster, T. M., and Steckel, L. E. (2013). Palmer
{Amaranth} ({Amaranthus} palmeri): {A Review}. \emph{Weed Technology}
27, 12--27.
doi:\href{https://doi.org/10.1614/WT-D-12-00113.1}{10.1614/WT-D-12-00113.1}.

\leavevmode\hypertarget{ref-webster2015}{}%
Webster, T. M., and Grey, T. L. (2015). Glyphosate-{Resistant Palmer
Amaranth} ({Amaranthus} palmeri) {Morphology}, {Growth}, and {Seed
Production} in {Georgia}. \emph{Weed Science} 63, 264--272.
doi:\href{https://doi.org/10.1614/WS-D-14-00051.1}{10.1614/WS-D-14-00051.1}.

\leavevmode\hypertarget{ref-wu2014}{}%
Wu, C., and Owen, M. D. K. (2014). When {Is} the {Best Time} to
{Emerge}: {Reproductive Phenology} and {Success} of {Natural Common
Waterhemp} ({Amaranthus} rudis) {Cohorts} in the {Midwest United
States}? \emph{Weed Science} 62, 107--117. Available at:
\url{http://www.jstor.org/stable/43700638}.

\leavevmode\hypertarget{ref-yacouby2020}{}%
Yacouby, R., and Axman, D. (2020). Probabilistic {Extension} of
{Precision}, {Recall}, and {F1 Score} for {More Thorough Evaluation} of
{Classification Models}. in \emph{Proceedings of the {First Workshop} on
{Evaluation} and {Comparison} of {NLP Systems}} ({Online}: {Association
for Computational Linguistics}), 79--91.
doi:\href{https://doi.org/10.18653/v1/2020.eval4nlp-1.9}{10.18653/v1/2020.eval4nlp-1.9}.

\leavevmode\hypertarget{ref-yu2021}{}%
Yu, E., Blair, S., Hardel, M., Chandler, M., Thiede, D., Cortilet, A.,
Gunsolus, J., and Becker, R. (2021). Timeline of {Palmer} amaranth
({Amaranthus} palmeri) invasion and eradication in {Minnesota}.
\emph{Weed Technology}, 1--31.
doi:\href{https://doi.org/10.1017/wet.2021.32}{10.1017/wet.2021.32}.

\end{CSLReferences}

\newpage

\hypertarget{figures}{%
\section*{Figures}\label{figures}}
\addcontentsline{toc}{section}{Figures}

\begin{figure}

{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}

}

\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 2}

}

\caption{Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit}\label{fig:Figure-2}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}

}

\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 4}

}

\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-4}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}

}

\caption{The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C). Black line represents the model estimation and shaded green the confidence intervals}\label{fig:Figure-5}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}

}

\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time. From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
\end{figure}

\end{document}